Risk reduction of SARS-CoV-2 transmission through disinfection of frequently touched surfaces. A Quantitative Microbial Risk Assessment.

### 2020-09-11 ES&T paper/Risk reduction of SARS-CoV-2 by disinfection of surfaces and hands.Rmd
Updating the code

\begin{tocentry}
Add image (TOC)
\end{tocentry}

@@

@@

@@

Frequently touched surfaces such as traffic light and train buttons are of risk concern due to their frequent use by multiple people.
SARS-CoV-2 contamination has been reported in public surfaces of an urban area in Brazil. Concentration of viruses [genome units /m2] was between 20 and 2990. The surfaces were bench, ground, handrail made out metal and concrete [@Abrahao2020].

## Surface disinfection

XX% reduction

@@

The proportion of viral particles emitted in a cough that is transferred to the hand when the hand is used to cover the mouth was estimated using a conical dispersion model  [@Nicas2006] -  the particles were assumed to be conically spread over a 80 angle, at a distance of 5cm. The transfer of viruses from the hand of the infected individual to the surface was estimated with: $Csurface=Chand*TEsh$, where $Chand$ is the concentration of viruses on the hand $[virus/cm^2]$, and TEsh $[%]$ is the transfer efficiency of viruses between the surface (assumed to be metal) and the hand. Once the surface was contaminated, the concentration of virus in the surface at any given time was calculated as follows $Csurf(t)= Csurf(0)$ $e^{-nt}$, where $Csurf(t)$ $[virus/cm^2]$ is the concentration of infective viruses at time $t$, $Csurf(0)$ $[virus/cm2^]$ is the initial concentration of virus in the surface, $t$ $[min]$ is the time lapsed after surface inoculation, and $n$ $[m^{-1}]$ is the decay rate of the virus in the selected surface (metal or plastic).

(Eq. \ref{eqn:example}).
\begin{equation}
  \frac{\mathrm{d}}{\mathrm{d}x} \, r^2 = 2r \label{eqn:example}
\end{equation}

\begin{equation}
  \begin{minipage}[c]{0.80\linewidth}
    \centering
    As illustrated here, the width of \\
    the minipage needs to allow some  \\
    space for the number to fit in to.
    %\includegraphics{graphic}
  \end{minipage}
  \label{eqn:graphic}
\end{equation}

# Experimental

\begin{acknowledgement}

The author thanks Sunil Dogga, Emmanuel Frousteyand, and Danilo Cuccato for their help with the model.

\end{acknowledgement}

\begin{suppinfo}

Experimental procedures and
characterization data for all new compounds. The class will
automatically add a sentence pointing to the information on-line:

\end{suppinfo}
## References

>>>>

\begin{tocentry}
\includegraphics{Risk-reduction-of-SARS-CoV-2-by-disinfection-of-surfaces-and-hands_files/figure-latex/TOC.pdf}
\end{tocentry}

@@

@Wilson2020 QMRA to evaluate the reduction in the risk of infection of SARS-CoV-2 for mask wearers.

@@

The proportion of viral particles emitted in a cough that is transferred to the hand when the hand is used to cover the mouth was estimated using a conical dispersion model  [@Nicas2009] -  the particles were assumed to be conically spread over a 80 angle, at a distance of 5cm. The transfer of viruses from the hand of the infected individual to the surface was estimated with: $Csurface=Chand*TEsh$, where $Chand$ is the concentration of viruses on the hand $[virus/cm^2]$, and TEsh $[%]$ is the transfer efficiency of viruses between the surface (assumed to be metal) and the hand. Once the surface was contaminated, the concentration of virus in the surface at any given time was calculated as follows $Csurf(t)= Csurf(0)$ $e^{-nt}$, where $Csurf(t)$ $[virus/cm^2]$ is the concentration of infective viruses at time $t$, $Csurf(0)$ $[virus/cm2^]$ is the initial concentration of virus in the surface, $t$ $[min]$ is the time lapsed after surface inoculation, and $n$ $[m^{-1}]$ is the decay rate of the virus in the selected surface (metal or plastic).

## Alternative way of presenting the equations
The equations can be presented as follows:

(Eq. \ref{eqn:Inactivation})
\begin{equation}
C(t)= C(0) e^{-nt} \label{eqn:Inactivation}
\end{equation}

(Eq. \ref{eqn:Transfer_sh})
\begin{equation}
C_{f_f} = C(t) TE_{sh} \label{eqn:Transfer_sh}
\end{equation}

(Eq. \ref{eqn:Dose})
\begin{equation}
D = C_{h} S_{hm} TE_{hm} \label{eqn:Dose}
\end{equation}

(Eq. \ref{eqn:DoseResponse})
\begin{equation}
P_{inf}=1 - e^{-kD} \label{eqn:DoseResponse}
\end{equation}

@@

Frequently touched surfaces such as and train buttons and  traffic light buttons are of risk concern due to their frequent use by multiple people. SARS-CoV-2 contamination has been reported in public surfaces of an urban area in Brazil. The concentration of viruses was between 20 and 2990  genome units per m$^2$. The surfaces were bench, ground, handrail made out metal, and concrete [@Abrahao2020]. In this article, we estimated the risk of contacting frequently touched surfaces as a function of the prevalence of infection within the population and evaluated the reduction of risk based on different disinfection methods.

“ How big the droplets need to be and how far away they can travel is still an open question"

@@

\begin{acknowledgement}
The author thanks Sunil Dogga, Emmanuel Frousteyand, and Danilo Cuccato for their help with the model. 
\end{acknowledgement}

\begin{suppinfo}
Experimental procedures and
characterization data for all new compounds. The class will
automatically add a sentence pointing to the information on-line:
\end{suppinfo}

## References

### 2020-09-11 ES&T paper/Risk-reduction-of-SARS-CoV-2-by-disinfection-of-surfaces-and-hands.tex
Updating the code

\begin{document}
\begin{abstract}
not exceeding 200 words A graphic for the Table of Contents (TOC) must
be supplied with each Letter This graphic can be no wider than 8 cm and
no taller than 4 cm and should be prepared following the specifications
under Artwork.
\end{abstract}
\begin{tocentry}
Add image (TOC)
\end{tocentry}

@@

@@

@@

Frequently touched surfaces such as traffic light and train buttons are
of risk concern due to their frequent use by multiple people. SARS-CoV-2
contamination has been reported in public surfaces of an urban area in
Brazil. Concentration of viruses {[}genome units /m2{]} was between 20
and 2990. The surfaces were bench, ground, handrail made out metal and
concrete.\textsuperscript{22}

\hypertarget{surface-disinfection}{%
\subsection{Surface disinfection}\label{surface-disinfection}}

XX\% reduction

@@

\hypertarget{maths}{%
\subsection{Math(s)}\label{maths}}

We modeled the risk of SARS-CoV-2 infection due to contact with
frequently touched buttons and estimated the effect that different
surface disinfection practices have on infection risk. The sequence of
events leading to the viral inoculation of the susceptible individual is
considered to be the following: 1) contamination event (infected person
coughing on the hand followed by a hand-to-surface contact), 2) virus
decay on the surface, 3) virus transfer from the surface to finger, and
4) virus transfer from finger to facial mucous membranes. The
stochastic-mechanistic model used herein is based on models described
elsewhere.\textsuperscript{17,23} Two scenarios were considered: risk of
infection with and without surface disinfection. Additionally, different
disinfection regimes were tested.

Viral loads in the saliva or sputum of symptomatic COVID-19 patients
within the first 14 days of symptom onset were used as model input. The
viral loads were measured using quantitative reverse-transcription
polymerase chain reaction (RT-qPCR).\textsuperscript{24--27} PCR data
does not distinguish between infectious and non-infectious virus. To
convert gene copies to infections virus, we used a ratio reported for
influenza viruses.\textsuperscript{28}

The proportion of viral particles emitted in a cough that is transferred
to the hand when the hand is used to cover the mouth was estimated using
a conical dispersion model\textsuperscript{{\textbf{???}}} - the
particles were assumed to be conically spread over a 80 angle, at a
distance of 5cm. The transfer of viruses from the hand of the infected
individual to the surface was estimated with: \(Csurface=Chand*TEsh\),
where \(Chand\) is the concentration of viruses on the hand
\([virus/cm^2]\), and TEsh \([%]
\) is the transfer efficiency of viruses between the surface (assumed to
be metal) and the hand. Once the surface was contaminated, the
concentration of virus in the surface at any given time was calculated
as follows \(Csurf(t)= Csurf(0)\) \(e^{-nt}\), where \(Csurf(t)\)
\([virus/cm^2]\) is the concentration of infective viruses at time
\(t\), \(Csurf(0)\) \([virus/cm2^]\) is the initial concentration of
virus in the surface, \(t\) \([min]\) is the time lapsed after surface
inoculation, and \(n\) \([m^{-1}]\) is the decay rate of the virus in
the selected surface (metal or plastic).

(Eq. \ref{eqn:example}). \begin{equation}
  \frac{\mathrm{d}}{\mathrm{d}x} \, r^2 = 2r \label{eqn:example}
\end{equation}

\begin{equation}
  \begin{minipage}[c]{0.80\linewidth}
    \centering
    As illustrated here, the width of \\
    the minipage needs to allow some  \\
    space for the number to fit in to.
    %\includegraphics{graphic}
  \end{minipage}
  \label{eqn:graphic}
\end{equation}

\hypertarget{experimental}{%
\section{Experimental}\label{experimental}}

\begin{acknowledgement}

The author thanks Sunil Dogga, Emmanuel Frousteyand, and Danilo Cuccato for their help with the model.

\end{acknowledgement}

\begin{suppinfo}

Experimental procedures and
characterization data for all new compounds. The class will
automatically add a sentence pointing to the information on-line:

\end{suppinfo}

@@

\leavevmode\hypertarget{ref-Abrahao2020}{}%
(22) Abrahao, J. S.; Pengo, L. S.; Rezende, I. M.; Rodrigues, R.;
Crispim, A. P. C.; Moura, C. C.; Mendonca, D. C.; Reis, E.; Souza, F.;
Oliveira, G. F. G. G. P. G.; Domingos, I. J. da S.; Boratto, P.; Silva,
P. H. B. e; Queiroz, V. F.; Silva, T. B. de S.; Oliveira, G. F. G. G. P.
G.; Alves, V. de S.; Alves, P. A.; Kroon, E. G.; Trindade, G. de S.;
Drumond, B. P.; Jônatas, A.; Abrahão, S.; Sacchetto, L.; Rezende, I. M.;
Rodrigues, R.; Paula, A.; Crispim, C.; Moura, C. C.; Correa, D.; Reis,
E.; Souza, F.; Fernanda, G.; Oliveira, G. F. G. G. P. G.; José, I.;
Domingos, S.; Boratto, P.; Henrique, P. \emph{medRxiv} \textbf{2020}.

\leavevmode\hypertarget{ref-Julian2018}{}%
(23) Julian, T. R.; Vithanage, H. S. K.; Chua, M. L.; Kuroda, M.; Pitol,
A. K.; Nguyen, P. H. L.; Canales, R. A.; Fujii, S.; Harada, H.
\emph{Science of The Total Environment} \textbf{2018}, \emph{635},
120--131.

\leavevmode\hypertarget{ref-Wolfel2020}{}%
(24) Wölfel, R.; Corman, V. M.; Guggemos, W.; Seilmaier, M.; Zange, S.;
Müller, M. A.; Niemeyer, D.; Jones, T. C.; Vollmar, P.; Rothe, C.;
Hoelscher, M.; Bleicker, T.; Brünink, S.; Schneider, J.; Ehmann, R.;
Zwirglmaier, K.; Drosten, C.; Wendtner, C. \emph{Nature} \textbf{2020},
1--14.

\leavevmode\hypertarget{ref-Pan2020}{}%
(25) Pan, Y.; Zhang, D.; Yang, P.; Poon, L. L. M.; Wang, Q. \emph{The
Lancet Infectious Diseases} \textbf{2020}, \emph{20} (4), 411--412.

\leavevmode\hypertarget{ref-Kim2020}{}%
(26) Kim, J. Y.; Ko, J. H.; Kim, Y.; Kim, Y. J.; Kim, J. M.; Chung, Y.
S.; Kim, H. M.; Han, M. G.; Kim, S. Y.; Chin, B. S. \emph{Journal of
Korean Medical Science} \textbf{2020}, \emph{35} (7), 1--7.

\leavevmode\hypertarget{ref-To2020}{}%
(27) To, K. K. W.; Tsang, O. T. Y.; Leung, W. S.; Tam, A. R.; Wu, T. C.;
Lung, D. C.; Yip, C. C. Y.; Cai, J. P.; Chan, J. M. C.; Chik, T. S. H.;
Lau, D. P. L.; Choi, C. Y. C.; Chen, L. L.; Chan, W. M.; Chan, K. H.;
Ip, J. D.; Ng, A. C. K.; Poon, R. W. S.; Luo, C. T.; Cheng, V. C. C.;
Chan, J. F. W.; Hung, I. F. N.; Chen, Z.; Chen, H.; Yuen, K. Y.
\emph{The Lancet Infectious Diseases} \textbf{2020}, \emph{20} (5),
565--574.

\leavevmode\hypertarget{ref-Ip2016}{}%
(28) Ip, D. K. M.; Lau, L. L. H.; Chan, K. H.; Fang, V. J.; Leung, G.
M.; Peiris, M. J. S.; Cowling, B. J. \emph{Clinical Infectious Diseases}
\textbf{2015}, \emph{62} (4), 431--437.
\end{document}

>>>>

\begin{document}
\begin{abstract}
not exceeding 200 words A graphic for the Table of Contents (TOC) must
be supplied with each Letter This graphic can be no wider than 8 cm and
no taller than 4 cm and should be prepared following the specifications
under Artwork.
\end{abstract}
\begin{tocentry}
\includegraphics{Risk-reduction-of-SARS-CoV-2-by-disinfection-of-surfaces-and-hands_files/figure-latex/TOC.pdf}
\end{tocentry}

@@

\textbf{???} QMRA to evaluate the reduction in the risk of infection of
SARS-CoV-2 for mask wearers.

@@

We modeled the risk of SARS-CoV-2 infection due to contact with
frequently touched buttons and estimated the effect that different
surface disinfection practices have on infection risk. The sequence of
events leading to the viral inoculation of the susceptible individual is
considered to be the following: 1) contamination event (infected person
coughing on the hand followed by a hand-to-surface contact), 2) virus
decay on the surface, 3) virus transfer from the surface to finger, and
4) virus transfer from finger to facial mucous membranes. The
stochastic-mechanistic model used herein is based on models described
elsewhere.\textsuperscript{17,22} Two scenarios were considered: risk of
infection with and without surface disinfection. Additionally, different
disinfection regimes were tested.

Viral loads in the saliva or sputum of symptomatic COVID-19 patients
within the first 14 days of symptom onset were used as model input. The
viral loads were measured using quantitative reverse-transcription
polymerase chain reaction (RT-qPCR).\textsuperscript{23--26} PCR data
does not distinguish between infectious and non-infectious virus. To
convert gene copies to infections virus, we used a ratio reported for
influenza viruses.\textsuperscript{27}

The proportion of viral particles emitted in a cough that is transferred
to the hand when the hand is used to cover the mouth was estimated using
a conical dispersion model\textsuperscript{28} - the particles were
assumed to be conically spread over a 80 angle, at a distance of 5cm.
The transfer of viruses from the hand of the infected individual to the
surface was estimated with: \(Csurface=Chand*TEsh\), where \(Chand\) is
the concentration of viruses on the hand \([virus/cm^2]\), and TEsh
\([%]
\) is the transfer efficiency of viruses between the surface (assumed to
be metal) and the hand. Once the surface was contaminated, the
concentration of virus in the surface at any given time was calculated
as follows \(Csurf(t)= Csurf(0)\) \(e^{-nt}\), where \(Csurf(t)\)
\([virus/cm^2]\) is the concentration of infective viruses at time
\(t\), \(Csurf(0)\) \([virus/cm2^]\) is the initial concentration of
virus in the surface, \(t\) \([min]\) is the time lapsed after surface
inoculation, and \(n\) \([m^{-1}]\) is the decay rate of the virus in
the selected surface (metal or plastic).

\hypertarget{alternative-way-of-presenting-the-equations}{%
\subsection{Alternative way of presenting the
equations}\label{alternative-way-of-presenting-the-equations}}

The equations can be presented as follows:

(Eq. \ref{eqn:Inactivation}) \begin{equation}
C(t)= C(0) e^{-nt} \label{eqn:Inactivation}
\end{equation}

(Eq. \ref{eqn:Transfer_sh}) \begin{equation}
C_{f_f} = C(t) TE_{sh} \label{eqn:Transfer_sh}
\end{equation}

(Eq. \ref{eqn:Dose}) \begin{equation}
D = C_{h} S_{hm} TE_{hm} \label{eqn:Dose}
\end{equation}

(Eq. \ref{eqn:DoseResponse}) \begin{equation}
P_{inf}=1 - e^{-kD} \label{eqn:DoseResponse}
\end{equation}

@@

Frequently touched surfaces such as and train buttons and traffic light
buttons are of risk concern due to their frequent use by multiple
people. SARS-CoV-2 contamination has been reported in public surfaces of
an urban area in Brazil. The concentration of viruses was between 20 and
2990 genome units per m\(^2\). The surfaces were bench, ground, handrail
made out metal, and concrete.\textsuperscript{29} In this article, we
estimated the risk of contacting frequently touched surfaces as a
function of the prevalence of infection within the population and
evaluated the reduction of risk based on different disinfection methods.

`` How big the droplets need to be and how far away they can travel is
still an open question"

@@

\begin{acknowledgement}
The author thanks Sunil Dogga, Emmanuel Frousteyand, and Danilo Cuccato for their help with the model. 
\end{acknowledgement}

\begin{suppinfo}
Experimental procedures and
characterization data for all new compounds. The class will
automatically add a sentence pointing to the information on-line:
\end{suppinfo}

@@

\leavevmode\hypertarget{ref-Julian2018}{}%
(22) Julian, T. R.; Vithanage, H. S. K.; Chua, M. L.; Kuroda, M.; Pitol,
A. K.; Nguyen, P. H. L.; Canales, R. A.; Fujii, S.; Harada, H.
\emph{Science of The Total Environment} \textbf{2018}, \emph{635},
120--131.

\leavevmode\hypertarget{ref-Wolfel2020}{}%
(23) Wölfel, R.; Corman, V. M.; Guggemos, W.; Seilmaier, M.; Zange, S.;
Müller, M. A.; Niemeyer, D.; Jones, T. C.; Vollmar, P.; Rothe, C.;
Hoelscher, M.; Bleicker, T.; Brünink, S.; Schneider, J.; Ehmann, R.;
Zwirglmaier, K.; Drosten, C.; Wendtner, C. \emph{Nature} \textbf{2020},
1--14.

\leavevmode\hypertarget{ref-Pan2020}{}%
(24) Pan, Y.; Zhang, D.; Yang, P.; Poon, L. L. M.; Wang, Q. \emph{The
Lancet Infectious Diseases} \textbf{2020}, \emph{20} (4), 411--412.

\leavevmode\hypertarget{ref-Kim2020}{}%
(25) Kim, J. Y.; Ko, J. H.; Kim, Y.; Kim, Y. J.; Kim, J. M.; Chung, Y.
S.; Kim, H. M.; Han, M. G.; Kim, S. Y.; Chin, B. S. \emph{Journal of
Korean Medical Science} \textbf{2020}, \emph{35} (7), 1--7.

\leavevmode\hypertarget{ref-To2020}{}%
(26) To, K. K. W.; Tsang, O. T. Y.; Leung, W. S.; Tam, A. R.; Wu, T. C.;
Lung, D. C.; Yip, C. C. Y.; Cai, J. P.; Chan, J. M. C.; Chik, T. S. H.;
Lau, D. P. L.; Choi, C. Y. C.; Chen, L. L.; Chan, W. M.; Chan, K. H.;
Ip, J. D.; Ng, A. C. K.; Poon, R. W. S.; Luo, C. T.; Cheng, V. C. C.;
Chan, J. F. W.; Hung, I. F. N.; Chen, Z.; Chen, H.; Yuen, K. Y.
\emph{The Lancet Infectious Diseases} \textbf{2020}, \emph{20} (5),
565--574.

\leavevmode\hypertarget{ref-Ip2016}{}%
(27) Ip, D. K. M.; Lau, L. L. H.; Chan, K. H.; Fang, V. J.; Leung, G.
M.; Peiris, M. J. S.; Cowling, B. J. \emph{Clinical Infectious Diseases}
\textbf{2015}, \emph{62} (4), 431--437.

\leavevmode\hypertarget{ref-Nicas2009}{}%
(28) Nicas, M.; Jones, R. M. \emph{Risk Analysis} \textbf{2009},
\emph{29} (9), 1292--1303.

\leavevmode\hypertarget{ref-Abrahao2020}{}%
(29) Abrahao, J. S.; Pengo, L. S.; Rezende, I. M.; Rodrigues, R.;
Crispim, A. P. C.; Moura, C. C.; Mendonca, D. C.; Reis, E.; Souza, F.;
Oliveira, G. F. G. G. P. G.; Domingos, I. J. da S.; Boratto, P.; Silva,
P. H. B. e; Queiroz, V. F.; Silva, T. B. de S.; Oliveira, G. F. G. G. P.
G.; Alves, V. de S.; Alves, P. A.; Kroon, E. G.; Trindade, G. de S.;
Drumond, B. P.; Jônatas, A.; Abrahão, S.; Sacchetto, L.; Rezende, I. M.;
Rodrigues, R.; Paula, A.; Crispim, C.; Moura, C. C.; Correa, D.; Reis,
E.; Souza, F.; Fernanda, G.; Oliveira, G. F. G. G. P. G.; José, I.;
Domingos, S.; Boratto, P.; Henrique, P. \emph{medRxiv} \textbf{2020}.
\end{document}
